The Shape Matching Element Method for Meshless Animation of NURBs Models

### 2021-01-17 paper/sections/methods.tex

\section{D-NURB Model}
\section{Shape Matching}
\section{Blending Polynomials}
\section{Raycasting Quadrature}
\section{Dynamic Simulation}
\subsection{Generalized Inertia}
\subsection{Generalized Forces}
\subsection{Equations of Motion}
This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.

@@

%% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
%% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$
%We have $\mathbf{x} = (x_0, x_1)$ and $\mathbf{X} = (X_0, X_1) $ as deformed and undeformed positions, respectively. \\
%\textbf{<insert Exposition about FEM >} \\

@@

%\subsection{Simulation on NURBs models}
%  \includegraphics[width=\linewidth]{figures/nurbs_surface}
%  \caption{The NURBs model uses control points, shown in grey, as the original degrees of freedom.}
%  \Description{A simple NURBS surface in $R^3$.}
%We used the D-NURBS model (ref) for simulating on NURBs surfaces, giving an approach to simulating directly on the degrees of freedom of parametric NURBs models. The equation for a NURBs curve takes the form:
%    \mathbf{x}(u) = \frac{\sum_{i=1}^{n} \mathbf{p}_i w_i B_{i,k}(u)}{ w_i B_{i,k}(u)}
%where $u$ is the parametric coordinate, $\mathbf{p}_i$ is the $i$-th control point with $n$ total control points. $B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$. NURBs curves represent a generalization of B-splines with the addition of the weight $w_i$ for each control point. In our case, we primarily work with NURBs surfaces, which are a generalization of the tensor-product of two B-splines:
%    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}{ w_{i,j} B_{i,k}(u)B_{j,l}(v)}
%where we now have two parameters $u,v$, and $m\cdot n$ control points. In D-NURBs, we see that $\mathbf{x}(u,v)$ is linear in the control points, $\mathbf{p}$, so we can rewrite the above equation as
%    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J}(u,v)\mathbf{p}.
%Here $\mathbf{J}$ is the Jacobian matrix that maps the control points to some world space position on the surface for some $u,v$ pair. <elaborate on the form of J>. Now with these generalized coordinates, we may perform dynamics directly on the degrees of freedom of our NURBs models. 

>>>>

\section{D-NURBS Model}
We provide a brief overview of the D-NURBS formulation presented in \cite{10.1145/176579.176580}. Our models in our method are composed of a collection of NURBS surfaces.

\subsection{D-NURBS Formulation}
A point on a NURBS surface is typically expressed in the following form
\begin{equation}
\label{eqn:nurbs_srf}
    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    {\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    \text{,}
\end{equation}
where where have a total of $nm$ control points $\mathbf{p}$ and weights $w$. Using the B-spline basis functions and the weighted control points, this formula gives a map from our parametric coordinates $u,v$ to their corresponding position on the surface in $\mathbb{R}^{3m}$. $B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$.

We would like our generalized coordinates to be the control points of the model so that the dynamics operates directly on the original representation. Therefore, we would like to rewrite equation \ref{eqn:nurbs_srf} in the form
\begin{equation}
    \mathbf{x}(u,v) = \mathbf{J}(u,v)\mathbf{q}
    \text{,}
\end{equation}
where $\mathbf{q} = \left[ \mathbf{p}_{1,1}^T, \dots, \mathbf{p}_{i,j}^T, \dots, \mathbf{p}_{n,m}^T \right]^T \in \mathbb{R}^{3nm}$ is the set of $nm$ control points \myworries{in the overview I just write $n$ control points, but here i'm writing $nm$} for a single NURBS surface arranged as a column vector. $\mathbf{J} \in  \mathbb{R}^{3 x 3nm}$ is the NURBS jacobian that maps $(u,v)$ coordinates to world positions in $\mathbb{R}^{3}$.

The NURBS jacobian, $\mathbf{J}$, is a matrix composed of horizontally concatenated $3x3$ blocks where the $(i,j)$-th block is $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ for the $(i,j)$-th control point. $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ is a diagonal $3x3$ matrix where each diagonal entry $N_{i,j}(\mathbf{u})$ takes the form

\begin{equation}
\label{eqn:jacobian_diagonal}
    N_{i,j}(u,v)
    = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    \text{.}
\end{equation}
Thus the NURBS jacobian for a single $(u,v)$ pair is written as
\begin{equation}
\label{eqn:uv_jacobian}
    \mathbf{J}(u,v)
    \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
    \right]
    \text{.}
\end{equation}

This formulation is a simplification of the full D-NURBS described in \cite{10.1145/176579.176580} the weights of the NURBS are fixed throughout the simulation. Permitting the dynamics to modify weights would likely improve the expressiveness of the kinematics, but we find fixing the weights affords sufficiently expressive results and improves performance. The performance improvement is a result of having fixed $N_{i,j}(u,v)$ entries over the course of the simulation. The full D-NURBS model requires rebuilding the jacobian and consequently the mass matrices at each time step.

\subsection{Generalized Coordinates for a NURBS model}
The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface. In our simulation we represent the entire boundary of the NURBS model, so we require samples along each surface. We emphasize that we only require an initial selection of $(u,v)$ coordinates across each surface and the dynamics of the simulation modifies the set of control points $\mathbf{q}$ to modify the world space maps $\mathbf{x}(u,v)=\mathbf{J}(u,v)\mathbf{q}$.

Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface. For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
\begin{equation}
\label{eqn:surface_jacobian}
    \mathbf{J}_i =
    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
    \text{.}
\end{equation}
Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$. Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
\begin{equation}
\mathbf{J} = \left( \begin{array}{ccccc}
\mathbf{J}_1 &  &  &  &  \\
 & \ddots &  &  &  \\
 &  & \mathbf{J}_i & &  \\
 &  &  & \ddots &  \\
 &  &  &  & \mathbf{J}_n \\
\end{array} \right)
\end{equation}
With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.

\section{Shape Matching}
\section{Blending Polynomials}
\section{Raycasting Quadrature}
\section{Dynamic Simulation}
\subsection{Generalized Inertia}
\subsection{Generalized Forces}
\subsection{Equations of Motion}
%This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.

% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$

@@

@@
